---
title: "Machine Learning Prediction Assignment Writeup"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Introduction
The goal of this project is to predict the manner in which participants performed barbell lifts, represented by the "classe" variable in the training dataset. We'll use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to build our prediction model.:

```{r , echo=TRUE}
# Load required libraries
library(caret)
library(randomForest)

# Load the data
training <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
testing <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")

# Remove near zero variance predictors
nzv <- nearZeroVar(training)
training <- training[, -nzv]
testing <- testing[, -nzv]

# Remove columns with more than 95% NA values
na_cols <- sapply(training, function(x) mean(is.na(x))) > 0.95
training <- training[, !na_cols]
testing <- testing[, !na_cols]

# Remove irrelevant columns (first 7 columns)
training <- training[, -(1:7)]
testing <- testing[, -(1:7)]

# Ensure 'classe' is a factor
training$classe <- as.factor(training$classe)
```

# Model Building
We'll use Random Forest for our prediction model due to its high accuracy and ability to handle complex interactions between features.

```{r, echo=TRUE}
# Set seed for reproducibility
set.seed(123)

# Split training data into training and validation sets
inTrain <- createDataPartition(training$classe, p = 0.7, list = FALSE)
train_set <- training[inTrain, ]
val_set <- training[-inTrain, ]

# Train Random Forest model
rf_model <- randomForest(classe ~ ., data = train_set, ntree = 500)
```

#MODEL EVALUATION
```{r, echo=TRUE}
# Make predictions on validation set
val_pred <- predict(rf_model, val_set)

# Calculate accuracy
accuracy <- confusionMatrix(val_pred, val_set$classe)$overall['Accuracy']
print(paste("Validation Accuracy:", round(accuracy, 4)))

# Calculate out-of-sample error estimate
oos_error <- 1 - accuracy
print(paste("Estimated Out-of-Sample Error:", round(oos_error, 4)))

```
#Predicting 20 test cases, output in console and submitted accordingly
```{r}
# Make predictions on the 20 test cases
test_pred <- predict(rf_model, testing)
#print(test_pred)
```

